---
title: Basic Inferential Data Analysis
Autor: Sadjad Zamani
output:
  html_document:
    keep_md: yes
  pdf_document: default

---
```{r setup, eval=TRUE, echo=FALSE, results="hide", message=FALSE}
library(tidyverse)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = "hide")
knitr::opts_chunk$set(warning = FALSE)
```

## Introduction
We want to analyze the The effect of vitamin C on Toothe Growth of guinea pigs.

For the final project of the course "[Statiscal Inference](https://www.coursera.org/learn/statistical-inference)" 
from Coursera.


### ToothGrowth
Let's first load the data and review it.

```{r toothgrowth, results="show"}
library(datasets)
data(ToothGrowth)
df <- as_tibble(ToothGrowth)
levels(df$supp) <- c("Orange Juice", "Ascorbic Acid")
summary(df)
```

So there are 60 observations of 3 variables:

- len: length of the tooth
- supp: type of supplement
- dose: dose of the supplement

```{r plot}
g <- ggplot(df) +
  aes(x = dose, y = len, color = supp) +
  geom_point()
g
```

According to the plot above, it's clear that as the dose increases, the length of the tooth increases.
Now I'm going to test two different hypothesis:

### Hypothesis 1
The "Orange Juice" and "Ascorbic Acid" have the same overall effect on the length of the tooth for all doses.

```{r hypothesis1, results="show"}
t.test(len ~ supp, data = df)
```

Since the p-value is above 0.05, we fail to reject the null hypothesis. So we can say that the "Orange Juice" and "Ascorbic Acid", "may" have the same overall effect on the length of the tooth for all doses.

### Hypothesis 2

Now let's check the null hypothesis of the "Orange Juice" and "Ascorbic Acid" having the same effect on the length of the tooth for dose equal to 0.5.

```{r hypothesis2, results="show"}
t.test(len ~ supp, data = subset(df, dose == 0.5))
```

Because the p-value is below 0.05, we reject the null hypothesis. So we can say that the "Orange Juice" and "Ascorbic Acid", "do not" have the same effect on the length of the tooth for dose equal to 0.5 and the "Orange Juice" has a stronger effect.

Keep in mind we could draw these conclusion also by looking at this plot:

```{r smooth,}
g +
  geom_smooth(method = lm, se = TRUE)
```

We can see there is no overlap between confidence intervals of the "Orange Juice" and "Ascorbic Acid" for dose equal to 0.5. And there is a strong overlap for dose equal to 2.