---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r include=FALSE}
# some global options
knitr::opts_chunk$set(warning = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(cache = TRUE)
knitr::opts_chunk$set(fig.path = "./figure/")
```

## Loading and preprocessing the data

Here I'm using the "readr" package to read the data which handles most of the required steps to clean the data.

```{r results="hide"}
library(readr)
library(tidyverse)
if (!file.exists("activity.zip")) {
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")
}
df <- read_csv("activity.zip")
```

Now let'a take a look at the data:

```{r}
dim(df)
head(df)
summary(df)
```

We can see there are two months of data, with some NA values in it. For now we'll leave them be.


## What is mean total number of steps taken per day?

Now we draw a plot of total number of steps taken each.

```{r daily,warning=FALSE}
dfDaily <- df %>%
  group_by(date) %>%
  summarise(total = sum(steps))

meanSteps <- mean(dfDaily$total, na.rm = TRUE)
medianSteps <- median(dfDaily$total, na.rm = TRUE)

ggplot(dfDaily) +
  geom_histogram(aes(x = total), bins = 40) +
  geom_vline(xintercept = meanSteps, color = "red") +
  geom_vline(xintercept = medianSteps, color = "blue")
```

We can see some missing lines in the graph.

Calculated mean is: `r as.integer(meanSteps)` and the meidan is: `r as.integer(medianSteps)`.

And we can see that the mean and median are very close.

## What is the average daily activity pattern?

```{r intervals}
df %>%
  group_by(interval) %>%
  summarise(mean = mean(steps, na.rm = T)) %>%
  ggplot(aes(x = interval, y = mean)) +
  geom_line() +
  geom_point(size = 0.5) +
  scale_x_continuous(n.breaks = 23) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5)) +
  stat_smooth(method = "lm", formula = y ~ 1, se = FALSE) +
  labs(x = "Time", y = "Mean steps", title = "Average steps in each interval")
```

And for finding the time interval where the most activity is in average:

```{r echo = TRUE}
averages <- df %>%
  group_by(interval) %>%
  summarise(mean = mean(steps, na.rm = T))

max <- averages[which.max(averages$mean), ]
max
```

And the code gices us the answer is: `r max$interval`.

## Imputing missing values

There are a total of `r sum(is.na(df$steps))` missing values in the data whcih is 
**`r as.integer(mean(is.na(df$steps))*100)`%** 
of the data.

For imputing the NA's I'll simply use the mean of that time interval.

```{r}
df2 <- df %>%
  group_by(interval) %>%
  mutate(steps = ifelse(is.na(steps), mean(steps, na.rm = T), steps))
```

Now we have df2 with no NA's.

```{r histogram,warning=FALSE}
dfDaily <- df2 %>%
  group_by(date) %>%
  summarise(total = sum(steps))

meanSteps <- mean(dfDaily$total, na.rm = TRUE)
medianSteps <- median(dfDaily$total, na.rm = TRUE)

ggplot(dfDaily) +
  geom_histogram(aes(x = total), bins = 40) +
  geom_vline(xintercept = meanSteps, color = "red") +
  geom_vline(xintercept = medianSteps, color = "blue")
```

Since we have replaced the NA's with the mean, we can see that the plot does not change so much. In fact, just some bars have a bit higher count and thats all.

## Are there differences in activity patterns between weekdays and weekends?

With this code we will add a column indicating weekday and weekends.
```{r}
df <- df %>%
  mutate(weekday = ifelse(strftime(date, "%a") == "Sat" | strftime(date, "%a") == "Sun", "weekend", "weekday")) %>%
  mutate(across(weekday, as.factor))

summary(df$weekday)
```

Now let's plot average step taken in each time interval for the two categories:

```{r weekend}
df %>%
  group_by(interval, weekday) %>%
  summarise(mean = mean(steps, na.rm = T)) %>%
  ggplot(aes(x = interval, y = mean)) +
  geom_line() +
  geom_point(size = 0.5) +
  facet_grid(rows = vars(weekday)) +
  scale_x_continuous(n.breaks = 23) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5)) +
  stat_smooth(method = "lm", formula = y ~ 1, se = FALSE) +
  labs(x = "Time", y = "Mean steps", title = "Average steps in each interval")
```
